SODA POP: SOft solving and Defect Aware Partial Ordering Planning

### 2016-04-15 article/article.tex
Finih ?

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            pdftitle={SODA POP: SOft solving and Defect Aware Partial Ordering Planning},
            pdfauthor={Paper ID\#232},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{listings}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}

\title{SODA POP: SOft solving and Defect Aware Partial Ordering Planning}
\author{Paper ID\#232}
\date{}
\usepackage{xcolor, listings}
\usepackage{latex/aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amsthm}

% We now define the sixteen \solarized{} colors.
\definecolor{solarized-base03} {RGB}{000, 043, 054}
\definecolor{solarized-base02} {RGB}{007, 054, 066}
\definecolor{solarized-base01} {RGB}{088, 110, 117}
\definecolor{solarized-base00} {RGB}{101, 123, 131}
\definecolor{solarized-base0}  {RGB}{131, 148, 150}
\definecolor{solarized-base1}  {RGB}{147, 161, 161}
\definecolor{solarized-base2}  {RGB}{238, 232, 213}
\definecolor{solarized-base3}  {RGB}{253, 246, 227}
\definecolor{solarized-yellow} {RGB}{181, 137, 000}
\definecolor{solarized-orange} {RGB}{203, 075, 022}
\definecolor{solarized-red}    {RGB}{220, 050, 047}
\definecolor{solarized-magenta}{RGB}{211, 054, 130}
\definecolor{solarized-violet} {RGB}{108, 113, 196}
\definecolor{solarized-blue}   {RGB}{038, 139, 210}
\definecolor{solarized-cyan}   {RGB}{042, 161, 152}
\definecolor{solarized-green}  {RGB}{133, 153, 000}

\newtheorem{definition}{Definition}

%%%%%% for algorithms
\errorcontextlines\maxdimen
\algrenewcommand\alglinenumber[1]{\tiny\color{solarized-base0} #1}

% begin vertical rule patch for algorithmicx (http://tex.stackexchange.com/questions/144840/vertical-loop-block-lines-in-algorithmicx-with-noend-option)
\makeatletter
% start with some helper code
% This is the vertical rule that is inserted
    \newcommand*{\algrule}[1][\algorithmicindent]{\makebox[#1][l]{\color{solarized-base0}\hspace*{.5em}\thealgruleextra\vrule height \thealgruleheight depth \thealgruledepth}}%
% its height and depth need to be adjustable
\newcommand*{\thealgruleextra}{}
\newcommand*{\thealgruleheight}{.75\baselineskip}
\newcommand*{\thealgruledepth}{.25\baselineskip}

\newcount\ALG@printindent@tempcnta
\def\ALG@printindent{%
    \ifnum \theALG@nested>0% is there anything to print
        \ifx\ALG@text\ALG@x@notext% is this an end group without any text?
            % do nothing
        \else
            \unskip
            \addvspace{-1pt}% FUDGE to make the rules line up
            % draw a rule for each indent level
            \ALG@printindent@tempcnta=1
            \loop
                \algrule[\csname ALG@ind@\the\ALG@printindent@tempcnta\endcsname]%
                \advance \ALG@printindent@tempcnta 1
            \ifnum \ALG@printindent@tempcnta<\numexpr\theALG@nested+1\relax% can't do <=, so add one to RHS and use < instead
            \repeat
        \fi
    \fi
    }%
\usepackage{etoolbox}
% the following line injects our new indent handling code in place of the default spacing
\patchcmd{\ALG@doentity}{\noindent\hskip\ALG@tlm}{\ALG@printindent}{}{\errmessage{failed to patch}}
\makeatother

% the required height and depth are set by measuring the content to be shown
% this means that the content is processed twice
\newbox\statebox
\newcommand{\myState}[1]{%
    \setbox\statebox=\vbox{#1}%
    \edef\thealgruleheight{\dimexpr \the\ht\statebox+1pt\relax}%
    \edef\thealgruledepth{\dimexpr \the\dp\statebox+1pt\relax}%
    \ifdim\thealgruleheight<.75\baselineskip
        \def\thealgruleheight{\dimexpr .75\baselineskip+1pt\relax}%
    \fi
    \ifdim\thealgruledepth<.25\baselineskip
        \def\thealgruledepth{\dimexpr .25\baselineskip+1pt\relax}%
    \fi
    %\showboxdepth=100
    %\showboxbreadth=100
    %\showbox\statebox
    \State #1%
    %\State \usebox\statebox
    %\State \unvbox\statebox
    %reset in case the next command is not wrapped in \myState
    \def\thealgruleheight{\dimexpr .75\baselineskip+1pt\relax}%
    \def\thealgruledepth{\dimexpr .25\baselineskip+1pt\relax}%
}
% end vertical rule patch for algorithmicx

%%%% For listings (code blocks)
\lstset{
    basicstyle={\footnotesize\ttfamily},
    numbers=left,
    keywordstyle=\color{solarized-orange}\bfseries,
%    identifierstyle=\color{solarized-base02},
    stringstyle=\color{solarized-cyan},
    commentstyle=\color{solarized-base1}\itshape,
    numberstyle=\tiny\color{solarized-base0},
    columns=flexible,
    stepnumber=1,
    numbersep=5pt,
    backgroundcolor=\color{solarized-base3},
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=2,
    captionpos=b,
    breaklines=true,
    breakatwhitespace=true,
    breakautoindent=true,
    escapeinside={\%*}{*)},
    linewidth=\textwidth,
    basewidth=0.5em,
}

%\setlength{\belowcaptionskip}{-2pt}
\setlength{\abovecaptionskip}{-12pt}

%%%% For links URL
\hypersetup{colorlinks=true, linkcolor=solarized-blue, citecolor=solarized-base1, filecolor=solarized-magenta, urlcolor=solarized-cyan}

% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\begin{document}
\maketitle
\begin{abstract}
As of recent years, automated planning domain has mainly focused on
performances and advances in state space planning to improve
scalability. That orientation shadows other less efficient ways of doing
like Partial Ordering Planning (POP) that has the advantage to be much
more flexible. This approach generates plans that can be refined easily
and can be used to achieve greater resilience, repairing capabilities
and soft resolution that applications based on online plan recognition
or decision-making can benefit from. This paper presents a set of
algorithms, named solving and Defect Aware Partial Ordering Planning
(SODA POP), that aims at targeting these objectives and maintain a high
plan quality. Our algorithms create proper offline plans for goals and
use an effective defect fixing algorithm to repair input plans even if
these plans are corrupted. SODA POP can also use healer actions and
links in order to always return a valid plan even in cases where the
problem is impossible to solve using problem derivation. Some relevant
properties of these algorithms are analyzed in this article and
experimental results show interesting performances for online planning
and repairing.
\end{abstract}

@@

For some time, Partial Order Planning (POP) has been the most popular
approach to planning resolution. This kind of algorithms are based on
\emph{least commitment strategy} on plan step ordering that allow
actions to be flexibly interleaved during execution {[}1{]}. Thus, the
way the search is made using flexible partial plan as a search space
allowed for more versatility for a wide variety of uses. As of more
recent years, new state space search approach and heuristics {[}2{]}
have been demonstrated to be more efficient than POP planners due to the
simplicity and relevance of state representations opposed to partial
plans {[}3{]}. This have made the search of performance the main axis of
research for planning.

While this goal is justified, it shadows other problems that some
practical applications cause. For example, the flexibility of Plan Space
Planning (PSP) is an obvious advantage for applications needing online
planning: plans can be repaired on the fly as new information and goals
are considered. The idea of using POP for online planning and repairing
plans instead of replanning from scratch is not new {[}4{]}, but has
never been paired with the resilience that some other cognitive
applications may need, especially when dealing with sensors noise in
input data. This resilience makes fixing errors easier than with an
external algorithm as the plan logic allows for context driven decision
on the way to repair the issues. For example, if an action becomes
unrelevant or incoherent, the flaw in the partial plan makes the issue
explicit and therefore easier to fix. Applications might sometimes
provide plans that can contain errors and imperfections that will
decrease significantly the efficiency of the computation and the quality
of the output plan. Additionnaly, these plans may become totally
unsolvable. This problem is to our knowledge not treated in planning of
all forms (state planning, PSP, and even constraint satisfaction
planning) as usually the aim is to find a solution relative to the
original plan (which makes sense). However, as we proceed a mechanism of
\emph{problem derivation} may be required. This will allow soft solving
of any problem regardless of its resolvability.

One of the applications that needs these improvements is plan
recognition with the particular use of off-the-shelf planners to infer
the pursued goal of an agent where online planning and resilience is
particularly important {[}5{]}. This method adds dummy actions that need
to be satisfied by modifying the goal, to ensure that the observed plan
is selected by the planner. A system that is able to repair plan will
ease such an application. Another application is decision-making in
dynamic environments. Indeed, having a plan that details all steps and
explicits the ones that are not possible, can help decide on the plan of
actions to take. These problems call for new ways to improve the answer
of a planner. These improvements must provide relevant plan information
pointing out exactly what needs to be done in order to make a planning
problem solvable, even in the case of obviously inconsistent input data.
This paper aims to solve this while preserving the advantages of PSP
algorithms (flexibility, easy fixing of plans, soundness and
completeness). We propose a new set of auxiliary algorithms that
combined forms our SOft solving and Defect Aware Partial Ordering
Planning (SODA POP) algorithm that targets those issues and allows
making POP algorithm more resilient, efficient and relevant. This is
done by pre-emptively computing proper plans for goals, by solving new
kinds of defects that input plans may exhibit, and by healing
compromised plan by deriving the initial problem with forged actions to
allow resolution. To explain this work we first introduce a few notions,
notations and specificities of existing POP algorithms. Then we present
and illustrate their limitations, categorizing the different defects
arising from the resilience problem and explaining our auxiliary
algorithms, their uses and properties. We compare the performance,
resilience and quality of POP and our solution.

\section{Definitions}\label{definitions}

First, we introduce some definitions and notations for mathematical
representation to present our model.

\subsection{Classical planning}\label{classical-planning}

\begin{definition}[Fluents]

A fluent is a property of the world. We note \(\lnot f\) the
complementary fluent of \(f\) meaning that \(f\) is true when
\(\lnot f\) is false and vice-versa.

@@

It is often represented by first order logical propositions but in this
paper we choose to focus on the algorithm and to represent fluents as
simple literals (fully instantiated) using \(\mathbb{Q}^*\), the set of
relative integers without \(0\), as the fluent domain. We use negative
integers to represent opposite fluents.

\begin{definition}[State]

A state is defined as a set of fluents. States can be additively
combined. We note
\(s_1 + s_2 = \left( s_1 \cup s_2 \right) - \left\{ f \middle| f \in s_1 \land \lnot f \in s_2 \right\}\)
such operation. It is the union of the fluents with an erasure of the
complementary ones.

@@

\begin{definition}[Action]

An action is a state operator. It is represented as a tuple
\(a = \langle pre, eff \rangle\) with \(pre\) and \(eff\) being sets of
fluents, respectively the preconditions and the effects of \(a\). An
action can be used only in a state that verifies its preconditions. We
note \(s \models a \Leftrightarrow pre(a) \subseteq s\) the verification
of an action \(a\) by a state \(s\).

@@

An action \(a\) can be functionally applied to a state \(s\) as follows:
\[a:= \substack{ \left\{ s \models a \middle| s \in S\right\} \to S\\
    a(s) \mapsto s + eff(a)}\] with \(S\) being the set of all states.
This means that an action adds all its effects to the state of the world
on application (by removing complementary fluents if needed). We
distinguish between two specific kinds of actions: actions with no
preconditions are synonymous to a state and those with empty effect are
called a goal.

\subsection{Plan Space Planning}\label{plan-space-planning}

\subsubsection{Problem}\label{problem}

We define a partial plan satisfaction problem as a tuple noted
\(P = \langle A, I, G, p \rangle\) with:

\begin{itemize}
\tightlist
\item
  \(I\) and \(G\) being the pseudo actions representing respectively the
  initial state and the goal.
\item
  \(p\) being a partial plan to refine.
\item
  \(A\) the set of all actions.
\end{itemize}

A \emph{partial plan} is a tuple \(p = \langle A_p, L\rangle\) where
\(A_p\) is a set of steps (actions) and \(L\) is a set of causal links
of the form \(a_i \xrightarrow{f} a_j\), such that
\(\{ a_i, a_j \} \subseteq A_p \land f \in eff(a_i) \cap pre(a_j)\) or
said otherwise, this causal link means that the fluent \(f\) is provided
by an effect of \(a_i\) to a precondition of \(a_j\). We include the
ordering constraints of PSP in the causal links. An ordering constraint
is noted \(a_i \to a_j\) and means that the plan means \(a_i\) as a step
that is prior to \(a_j\) without specific cause (usually because of
threat resolution). We note the order relation over \(A\) for an action
\(a_i\) that is prior to the action \(a_j\) following all the causal
links and ordering constraints \(a_i \succ a_j\). We can consider it as
a graph formed by step as vertices and causal links as edges.

The figure \ref{fig:legend} details how the elements of partial plans
are represented in the rest of this paper.

\begin{figure}[htbp]
\centering
\includegraphics{graphics/legend.pdf}
\caption{Global legend for partial plans\label{fig:legend}}
\end{figure}

In this figure the fluent \(6\) is a precondition of the action \(a\)
and also an effect. The causal link is providing the fluent \(6\). The
dotted line is an ordering constraint specifying which action should be
placed before another. \(I\) and \(G\) are respectively the initial and
goal step.

\subsubsection{Flaws}\label{flaws}

When refining a partial plan, we need to fix flaws. Those could be
present in the input or created by the refining process. Flaws can
either be unsatisfied subgoals or threats to causal links.

@@

A subgoal \(s\) is a precondition of an action \(a_s \in A_p\) with
\(s \in pre(a_s)\) that isn't satisfied by any causal link. We can note
a subgoal as:
\[a_i \xrightarrow{s} a_s \notin L \mid \{ a_i, a_s \} \subseteq A_p \]

@@

A step \(a_t\) is said to threaten a causal link
\(a_i \xrightarrow{t} a_j\) if and only if
\[\neg t \in eff(a_t) \land a_i \succ a_t \succ a_j \models L\]

@@

In other words, the action has a complementary effect that can be
inserted between two actions needing this fluent while being consistent
with the ordering constraint in \(L\).

\begin{definition}[Resolvers]

Resolvers are a set of actions and causal links
\(r = \langle A_r , L_r \rangle\) that fixes flaws.

\begin{itemize}
\tightlist
\item
  A resolver for a subgoal is an action \(a_r \in A\) that has \(s\) as
  an effect \(s \in eff(a_r)\) inserted along with a causal link noted
  \(a_r \xrightarrow{s} a_s\).
\item
  The usual resolvers for a threat are either \(a_t \to a_i\) or
  \(a_j \to a_t\) which are called respectively promotion and demotion
  links.
\end{itemize}

@@

\begin{definition}[Solution]

The solution of a PSP problem is a valid partial plan that respects the
specification of said problem (only using actions in \(A\) and having
the correct initial and goal step).

@@

\begin{definition}[Consistency]

A partial plan is consistent if it contains no ordering cycles. The
directed graph formed by step as vertices and causal links as edges
isn't cyclical. This is important to guarantee the soundness.

>>>>

@@

Until the end of the 90s, Plan-Space Planning (PSP) was generally
preferred by the automated planning community. Its early commitment,
expressiveness, and flexibility were clear advantages over State-Space
Planning (SSP). However, more recently, drastic improvements in state
search planning were made possible by advanced and efficient heuristics.
This allowed those planners to scale up more efficiently than plan-space
search ones, notably thanks to approaches like GraphPlan
{[}\protect\hyperlink{ref-blumux5ffastux5f1997}{1}{]}, fast-forward
{[}\protect\hyperlink{ref-hoffmannux5fffux5f2001}{2}{]}, LAMA
{[}\protect\hyperlink{ref-richterux5flamaux5f2011}{3}{]} and Fast
Downward Stone Soup
{[}\protect\hyperlink{ref-rogerux5ffastux5f2014}{4}{]}. This evolution
led to a preference for performances upon other aspects of the problem
of automated planning. However, some of these aspects can be more easily
addressed in Partial Order Planning (POP), also known as Partial Order
Causal Link (POCL) planning. For example POP, can take advantage of
least commitment
{[}\protect\hyperlink{ref-mccluskeyux5fengineeringux5f1997}{5}{]} that
offers more flexibility with a partial plan that describes only the
necessary order of the actions and not a fixed sequence of steps. POP
has been proven to be well suited for multi-agent planning
{[}\protect\hyperlink{ref-kvarnstromux5fplanningux5f2011}{6}{]} and
temporal planning
{[}\protect\hyperlink{ref-bentonux5ftemporalux5f2012}{7}{]} . These
advantages made UCPOP
{[}\protect\hyperlink{ref-penberthyux5fucpopux5f1992}{8}{]} one of the
preferred POP planners of its time with works made to port some of its
characteristics into state-based planning
{[}\protect\hyperlink{ref-gazenux5fcombiningux5f1997}{9}{]}.

Related works already tried to explore new ideas to make POP an
attractive alternative to regular state-based planners like the
appropriately named ``Reviving partial order planning''
{[}\protect\hyperlink{ref-nguyenux5frevivingux5f2001}{10}{]} and VHPOP
{[}\protect\hyperlink{ref-younesux5fvhpopux5f2003}{11}{]}. More recent
efforts
{[}\protect\hyperlink{ref-colesux5fforwardchainingux5f2010}{12}{]},
{[}\protect\hyperlink{ref-sapenaux5fcombiningux5f2014}{13}{]} are trying
to adapt the powerful heuristics from state-based planning to POP's
approach. An interesting approach of these last efforts is found in
{[}\protect\hyperlink{ref-shekharux5flearningux5f2016}{14}{]} with
meta-heuristics based on offline training on the domain. However, we
clearly note that only a few papers lay the emphasis upon plan quality
using POP {[}\protect\hyperlink{ref-ambiteux5fplanningux5f1997}{15}{]},
{[}\protect\hyperlink{ref-estlinux5flearningux5f1997}{16}{]}.

The present paper lays the base for our project for an intelligent
robotic system that aims to use inverted planning for plan inference in
order to help dependent persons to accomplish tasks. This project is
based on the works of Ramirez et al.
{[}\protect\hyperlink{ref-ramirezux5fplanux5f2009}{17}{]} on inverted
planning for plan inference. This context led us to seek ways to improve
POP with better refining techniques and resolver selection. Since we
need to handle data derived from limited robotic sensors, we need a way
for the planner to be able to be resilient to misleading input plans.
Another aspect of this work lies in the fact that the final system will
need to compute online planning with a feed of observational data. In
order to achieve this we need a planner that can:

\begin{itemize}
\tightlist
\item
  repair and optimize existing plans,
\item
  perform online planning efficiently
\item
  retain performances on complex but medium sized problems.
\end{itemize}

Classical POP algorithms don't meet most of these criteria but can be
enhanced to fit the first one. Usually, POP algorithms take a problem as
an input and use a loop or a recursive function to refine the plan into
a solution. We can't simply use the refining recursive function in order
to be able to use our existing partial plan. This causes multiples side
effects if the input plan is suboptimal. \emph{Our view} on the problem
diverges from other works: PSP is a very different approach compared to
SSP. It is usually more computationally expensive than modern state
space planners but brings several advantages. We want to make the most
of them instead of trying to change POP's fundamental nature. That view
is at the core of our model: we use the refining and least commitment
abilities of POP in order to improve online performances and quality. In
order to achieve this, we start by computing a \emph{proper plan} that
is computed offline with the input of the domain. The explanation of the
way this notion is defined and used is given in \cref{sec:properplan} of
the present paper. Using existing partial plans as input leads to
several issues, mainly with new flaw types that aren't treated in
classical POP. This is why we focus the \cref{sec:negative} of our paper
on plan quality and negative refinements. We, therefore, introduce new
negative flaws and resolvers that aim to fix and optimize the plan: the
\emph{alternative} and the \emph{orphan} flaws. Side effects of negative
flaws and resolvers can lead to conflicts. In order to avoid them and
enhance performances and quality, the algorithm needs resolver and flaw
selection mechanisms that are explained in the \cref{sec:selection}. All
these mechanisms are part of our aLternative Optimization with partiaL
pLan Injection Partial Ordered Planning (LOLLIPOP) algorithm presented
in details in \cref{sec:algorithm}. We prove that the use of these new
mechanisms keeps LOLLIPOP sound and complete in \cref{sec:analysis}.
Experimental results and benchmarks are presented and explained in the
\cref{sec:results} of this paper.

\section{Partial Order Planning
Preliminaries}\label{partial-order-planning-preliminaries}

\subsection{Notation}\label{notation}

For the rest of this paper, we will use the notation defined in
\cref{tbl:symbols}. In order to make some writings more concise, we use
the symbol \(\pm\) to signify that there is a notation for the positive
and negative symbols but the current formula works regardless the sign.
All related notions will be defined later.

\begin{table}\footnotesize
\centering

\caption{\label{tbl:symbols}Most used symbols in the paper. }

\begin{tabular}{@{}ll@{}}
\toprule

Symbol & Description \\\midrule

\(pre(o)\), \(eff(o)\) & Preconditions and effects of the operator
\(o\) \\
\(\Delta\) & Planning domain \\
\(\Pi\) & Planning problem \\
\(T.x\) & Access element \(x\) of tuple \(T\) \\
\(l_{\rightarrow}\), \(l_{\leftarrow}\) & Source and target of the
causal link \(l\) \\
\(o_1 \succ o_2\) & Precedence operator (\(o_1\) precedes \(o_2\)) \\
\(O^P\) & Proper plan of the set of operators \(O\) \\
\(p.d^\pm(o)\) & Outgoing and incoming degrees of \(o\) \\
\(o.d^\pm\) & Proper degrees of \(o\) (\(|pre(o)|\) and \(|eff(o)|\)) \\
\(p.L^\pm(o)\) & Outgoing and incoming causal links of \(o\) \\
\(C(p)\) & Set of cycles in partial plan \(p\) \\
\(C_p(o)\) & Set of cycles in \(p\), \(o\) is part of \\
\(SC_p(o)\) & \(\{o\}\) if \(o\) has a self cycle in \(p\),
\(\emptyset\) otherwise \\
\(F^\pm(p)\) & Set of flaws in \(p\) \\
\(r(f)\) & Resolvers of the flaw \(f\) \\
\(f.n\) & Needer of the flaw \(f\) \\
\(f(p)\) & Application of the flaw \(f\) on plan \(p\) \\
\(fs(p)\) & Full support of \(p\) \\
\(p \models \Pi\) & The partial plan \(p\) is a valid solution of
\(\Pi\) \\

\bottomrule
\end{tabular}

\end{table}

\subsection{Basic Definitions}\label{basic-definitions}

Planning systems need a representation of its fluents and operators. Our
framework is based on a classical domain definition. This is defined in
the following definition from
{[}\protect\hyperlink{ref-gobelbeckerux5fcomingux5f2010}{18}{]}.

\begin{definition}[Domain]\label{def:domain}

We define our planning domain as a tuple
\(\Delta = \langle T, C, P, F, O \rangle\) where

\begin{itemize}
\tightlist
\item
  \(T\) are the \textbf{types},
\item
  \(C\) is the set of \textbf{domain constants},
\item
  \(P\) is the set of \textbf{properties} with their arities and typing
  signatures,
\item
  \(F\) represents the set of \textbf{fluents} defined as potential
  equations over the terms of the domain,
\item
  \(O\) is the set of optionally parameterized \textbf{operators} with
  preconditions and effects.
\end{itemize}

@@

Along with a domain, every planner needs a problem representation. For
this, we use the classical problem representation with some special
additions.

\begin{definition}[Problem]\label{def:problem}

The planning problem is defined as a tuple
\(\Pi = \langle \Delta, C_\Pi, I, G, p\rangle\) where

\begin{itemize}
\tightlist
\item
  \(\Delta\) is a planning domain,
\item
  \(C_\Pi\) is the set of \textbf{problem constants} disjoint from the
  domain constraints \(C\),
\item
  \(I\) is the \textbf{initial state},
\item
  \(G\) is the \textbf{goal},
\item
  \(p\) is a given \textbf{partial plan}.
\end{itemize}

@@

The framework uses the \emph{closed world assumption} in which all
predicates and properties that aren't defined in the initial step are
assumed false or don't have a value.

\Cref{fig:example} portrays an example of a planning domain and problem
that we use as a guideline throughout the article.

\begin{figure}[htbp]
\centering
\includegraphics{graphics/example.pdf}
\caption{\label{fig:example}Example domain and problem featuring a robot
that aims to fetch a lollipop in a locked kitchen. The operator
\lstinline!go! is used for movable objects (such as the robot) to move
to another room. The \lstinline!grab! operator is used by grabbers to
hold objects and the \lstinline!unlock! operator is used to open a door
when the robot holds the key.}
\end{figure}

In order to simplify this framework, we need to introduce some
differences from the classical partial plan representation. First, the
partial plan is a part of the problem tuple as it is a needed input of
the LOLLIPOP algorithm.

We define a partial plan as a tuple \(\langle S, L, B\rangle\) with
\(S\) the set of \textbf{steps} (semi or fully instantiated operators
also called \emph{actions}), \(L\) the set of \textbf{causal links}, and
\(B\) the set of \textbf{binding constraints}.

@@

Second we factorize the set of \emph{ordering constraints}, used in
classical representations, as being part of the causal links. Indeed,
causal links are always supported by an ordering constraint. The only
case where bare ordering constraints are needed is in threats. We
represent them with \textbf{bare causal links}. These are stored as
causal links without bearing any fluents. Causal links can be
represented by their bore fluents called \emph{causes}. We note
\(f \in l\) the fact that a causal link \(l\) bears the fluent \(f\)
(bare causal links are noted \(l = \emptyset\)). That allows us to
introduce the \textbf{precedence operator} noted \(a_i \succ a_j\) with
\(a_i, a_j \in S\) iff there is a path of causal links that connects
\(a_i\) to \(a_j\). We call \(a_i\) \emph{anterior} to \(a_j\).

A specificity of Partial Order Planning is that it fixes flaws in a
partial plan in order to refine it into a valid plan that is a solution
to the given problem. Next, we define the classical flaws in our
framework.

@@

A flaw in a partial plan, called subgoal, is a missing causal link
required to satisfy a precondition of a step. We can note a subgoal as:
\(a_p \xrightarrow{s} a_n \notin L \mid \{ a_p, a_n \} \subseteq S\)
with \(a_n\) called the \textbf{needer} and \(a_p\) an eventual
\textbf{provider} of the fluent \(s\). This fluent is called \emph{open
condition} or \textbf{proper fluent} of the subgoal.

@@

A flaw in a partial plan called threat consists of having an effect of a
step that can be inserted between two actions with a causal link that is
threatened by the said effect. We say that a step \(a_b\) is threatening
a causal link \(a_p \xrightarrow{t} a_n\) iff
\(a_b \neq a_p \neq a_n \land \neg t \in eff(a_b) \land a_p \succ a_b \succ a_n\)
with \(a_b\) being the \textbf{breaker}, \(a_n\) the \emph{needer} and
\(a_p\) a \emph{provider} of the \emph{proper fluent} \(t\).

@@

Flaws are fixed via the application of a resolver to the plan. A flaw
can have several resolvers that match its needs.

\begin{definition}[Resolvers]\label{def:resolver}

A resolver is a potential causal link defined as a tuple
\(r = \langle a_s, a_t, f\rangle\) with:

\begin{itemize}
\tightlist
\item
  \(a_s, a_t \in S\) being the source and the target of the resolver,
\item
  \(f\) being the considered fluent.
\end{itemize}

@@

For classical flaws, the resolvers are simple to find. For a
\emph{subgoal} the resolvers are sets of the potential causal links
between a possible provider of the proper fluent and the needer. To
solve a \emph{threat} there are mainly two resolvers: a causal link
between the needer and the breaker called \textbf{demotion} or a causal
link between the breaker and the provider called \textbf{promotion}.

Once the resolver is applied, another important step is needed in order
to be able to keep refining. The algorithm needs to take into account
the \textbf{side effects} the application of the resolver had on the
partial plan. Side effects are searched by type.

\begin{definition}[Side effects]\label{def:sideeffect}

Flaws that arise because of the application of a resolver on the partial
plan are called causal side effects or \emph{related flaws}. They are
caused by an action \(a_t\) called the \textbf{trouble maker} of a
resolver. This action is the \emph{source} of the resolver applied onto
the plan.

@@

We can derive this definition for subgoals and threats:

\begin{itemize}
\tightlist
\item
  \textbf{Related Subgoals} are all the open conditions inserted with
  the \emph{trouble maker}. The subgoals are often searched using the
  preconditions of the trouble maker and added when no causal links
  satisfy them.
\item
  \textbf{Related Threats} are the causal links threatened by the
  insertion of the \emph{trouble maker}. They are added when there is no
  causal path that prevents the action to interfere with the threatened
  causal link.
\end{itemize}

\begin{figure}[htbp]
\centering
\includegraphics{graphics/partialplan1.pdf}
\caption{\label{fig:partialplan1}Example of a partial plan occurring
during the computation of POP on the previous example domain illustrated
by \cref{fig:example}. Full arrows are existing causal links. The dotted
arrow and operator depict the current resolver.}
\end{figure}

In the partial plan presented in \cref{fig:partialplan1}, we consider
that a resolver providing the fluent \lstinline!robot holds key! is
considered. This resolver will introduce the open conditions
\lstinline!robot holds nothing, key room _room, robot room _room! since
it just introduced this instantiation of the \lstinline!grab! operator
in the partial plan. These will trigger the insertion of a subgoal that
will have this new \lstinline!grab! operator as needer. The potentially
related threat of this resolver is that the effect
\lstinline!robot holds key! might threaten the link between the existing
\lstinline!unlock! and \lstinline!grab! steps, but won't be considered
since there are no way the new step can be inserted after
\lstinline!unlock!.

\subsection{Classical POP Algorithm}\label{classical-pop-algorithm}

The classical POP algorithm is pretty straight forward: it starts with a
simple partial plan and refines its \emph{flaws} until they are all
resolved to make the found plan a solution of the problem.

\begin{algorithm}\caption{Classical Partial Order Planning}\label{alg:pop}\begin{algorithmic}[1]

\footnotesize
\Function{pop}{Queue of Flaws $a$, Problem $\Pi$}
\State \Call{populate}{$a$, $\Pi$}
\Comment{Populate agenda only on first call} \If{$\Pi.G = \emptyset$}
\Comment{Goal is empty, default solution is provided}
\State \(\Pi.p.L \gets (I \rightarrow G)\) \label{line:emptygoal} \EndIf
 \If{$a = \emptyset$} \State \Return Success
\Comment{Stop all recursion} \EndIf
 \State Flaw \(f \gets\) \Call{choose}{$a$} \label{line:flawselection}
\Comment{Non deterministic choice} \State Resolvers \(R \gets\)
\Call{resolvers}{$f$, $\Pi$} \ForAll{$r \in R$}
\Comment{Non deterministic choice operator}
\State \Call{apply}{$r$, $\Pi.p$} \label{line:resolverapplication}
\Comment{Apply resolver to partial plan} \State SideEffects \(s \gets\)
\Call{sideEffect}{$r$} \label{line:sideeffectapplication}
\State \Call{apply}{$s$, $a$} \Comment{Side effects of the resolver}
\If{\protect\Call{pop}{$a$, $\Pi$} = Success}
\Comment{Refining recursively} \State \Return Success \EndIf
 \State \Call{revert}{$r$, $\Pi.p$}
\Comment{Failure, undo resolver insertion}
\State \Call{revert}{$s$, $a$}
\Comment{Failure, undo side effects application} \EndFor
 \State \Return Failure
\Comment{Revert to last non deterministic choice of resolver}
\EndFunction

The algorithm~\ref{alg:pop} is inspired by
{[}\protect\hyperlink{ref-ghallabux5fautomatedux5f2004}{19}{]}. This POP
implementation uses an agenda of flaws that is efficiently updated after
each refinement of the plan. At each iteration, a flaw is selected and
removed from the agenda (line~\ref{line:flawselection}). A resolver for
this flaw is then selected and applied
(line~\ref{line:resolverapplication}). If all resolvers cause failures,
the algorithm backtracks to the last resolver selection to try another
one. The algorithm terminates when no more resolver fits a flaw
(\lstinline!Failure!) or when all flaws have been fixed
(\lstinline!Success!).

This standard implementation has several limitations. First, it can
easily make poor choices that will lead to excessive backtracking. It
also can't undo redundant or non-optimal links if they don't lead to
backtracking. To explain these limitations, we use the example described
in \cref{fig:example} where a robot must fetch a lollipop in a locked
room. This problem is solvable by regular POP algorithms. However, we
can have some cases where small changes in POP's inputs can cause a lot
of unnecessary back-tracking. For example, if we add a new action called
\lstinline!dig_through_wall! that has as effect to be in the desired
room but that requires a \emph{jackhammer}, the algorithm will simply
need more backtracking. The effects could be worse if obtaining a
jackhammer would require numerous steps (for example needing to build
it). This problem can be solved most of the time using simple flaw
selection mechanisms. The other limitation arises when the plan has been
modified. This can arise in the context of dynamic environments of
online planning applications. When POP is modified to use input plans
{[}\protect\hyperlink{ref-sebastiaux5fgraph-basedux5f2000}{20}{]}, these
partial plans are carefully designed not to infringe the properties
required by POP to operate. In our case, the plan can contain misleading
information and this can cause a variety of new problems that can only
be fixed using new refinements methods.

\section{LOLLIPOP's Approach}\label{lollipops-approach}

LOLLIPOP makes use of proper plans in order to ease the initial
backchaining of POP and to drive the flaw and resolver selection,
including the new negative flaws we introduced for online planning
refinements.

\subsection{Proper Plan Generation}\label{sec:properplan}
